---
title: "CHM Group Uncertainty Project"
author: "Kristin Braziunas"
date: "June 22, 2016"
output: html_document
---

## About

For this project we will be looking at LiDAR-generated Canopy Height Models and looking at uncertainty compared to field measurements of plot-level and tree-level characteristics.

For this project, I am going to be working with data from NEON Domain 01, Harvard Forest Field Site (HARV). I downloaded my data from NEON's figshare site.

Specific characteristics we will be looking at include...
*Fill this part in

## First, I'm going to load the libraries I need to work with the data.

Also, added additional setup steps in this code chunk.

```{r load-libraries }

library(raster)
library(rgdal)
library(dplyr)

# also set stringsasFactors to false
options(stringsAsFactors = FALSE)

```

## Next, I'm going to load in the canopy height model

```{r import-chm }

site_chm <- raster("../NEONdata/D01-Massachusetts/HARV/2014/lidar/HARV_lidarCHM.tif")

# take a look at it
plot(site_chm, main="Canopy Height Model\nHARV Field Site")
hist(site_chm, main="CHM Histogram\nHARV Field Site")

# set 0 values to NA
site_chm[site_chm == 0] <- NA

# take another look at the histogram
hist(site_chm, main="CHM Histogram\nHARV Field Site",
     xlab="Height (m)")

```

## Then, I'm going to load in the insitu data

```{r import-insitu }

site_insitu <- read.csv("../NEONdata/D01-Massachusetts/HARV/2014/insitu/veg_structure/D01_2012_HARV_vegStr.csv", 
                        header=TRUE, sep=",")

# let's make sure the data came in okay
head(site_insitu)  # lots of NA values

# look at a histogram of the heights
hist(site_insitu$stem_height, main="In Situ Histogram\nHARV Field Site",
     xlab="Height (m)")  

```

In the insitu data, it looks like there's a lot more shorter vegetation than was captured in the CHM.

## Load in the plot locations

```{r plot-locations }

# loading using readOGR

site_plots <- readOGR("../NEONdata/D01-Massachusetts/HARV/vector_data/",
                      "HARV_PlotCentroids")

# look at the plot locations
plot(site_plots)

# overlay plot locations on CHM
plot(site_chm, main="CHM with In Situ Locations\nHARV Field Site")
plot(site_plots, add=TRUE)

```

# Following along with NEON tutorial on creating field data polygons from centroids.
**Add link here**

## Set up plot dimensions

I'm going to set up square plots around the plot centroid. This process involves setting a radius, which will be half the length of one side of the square.

This is based on the assumption that plots are oriented north and not rotated.

```{r set-plot-dimensions }

# first set radius

radius <- 20  # in meters

# then define the plot boundaries using the radius
yPlus <- site_plots$Y + radius
xPlus <- site_plots$X + radius
yMinus <- site_plots$Y - radius
xMinus <- site_plots$X - radius

```

## Extract Plot IDs from centroids

```{r extract-plot-id }

ID <- site_plots$plotID

# let's look at our plot IDs
ID

```

## Now I'm going to calculate square polygon coordinates for each plot centroid

```{r polygon-coordinates }

# here I am just making a list of the coordinates of each corner of the square
# will follow x, y, x, y, etc.
# later we will turn this into the 4 corners of the square the bounds the plot
# this requires 5 corners to draw the square - why are there 6?

square <- cbind(xMinus, yPlus, 
                xPlus, yPlus, 
                xPlus, yMinus, 
                xMinus, yMinus, 
                xMinus, yPlus,
                xMinus, yPlus)

```

## Create spatial polygons using the coordinates

```{r create-spatial-polygons }

# Create a function to do this
polys <- SpatialPolygons(mapply(function(poly, id) {
  xy <- matrix(poly, ncol=2, byrow=TRUE)  # take a list and create a matrix
  Polygons(list(Polygon(xy)), ID=id)
}, split(square, row(square)), ID),proj4string=CRS(as.character("+proj=utm +zone=18 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")))

```

## Create shapefile

This is an important step to extract the data.

```{r create-shapefile }

polys.df <- SpatialPolygonsDataFrame(polys, data.frame(id=ID, row.names=ID))

```

## Plot this with our CHM

```{r plot-square-buffers }

plot(site_chm, main="CHM with square in situ plots\nHARV Field Site")
plot(polys.df, add=TRUE)

```




## Use square buffers to extract information from LiDAR data

```{r extract-info-from-lidar }

# extracting max height from CHM at field data sites
chm_maxheight <- extract(site_chm, polys.df,
                       fun=max, sp=TRUE,  # sp for return spatial object
                       stringsasFactors=FALSE)

# extract mean height from CHM at field data sites
chm_meanheight <- extract(site_chm, polys.df,
                       fun=mean, sp=TRUE,  # sp for return spatial object
                       stringsasFactors=FALSE)

# check outputs
chm_maxheight@data
chm_meanheight@data
unique(site_insitu$plotid)

```

## Extract information from insitu data

```{r extract-info-from-insitu, eval = FALSE}

# extract max ht by plot from insitu data with dplyr
insitu_maxheight <- site_insitu %>% 
  group_by(plotid) %>%
  summarise(max(stem_height))

names(insitu_maxheight) <- c("plotid","max_height")

hist(insitu_maxheight$max_height)
hist(chm_maxheight@data$HARV_lidarCHM)

# quick look at data summary
summary(fieldmaxht.sjer)
summary(height.sjer@data$SJER_lidarCHM)

# update names
names(fieldmaxht.sjer) <- c("plotid","field_maxht")


```

