---
title: "group_uncertainty"
author: "Megan Cattau"
date: "June 22, 2016"
output: html_document
---

We want to compare data that we collected on the ground with LiDAR-derived raster data
We need to extract data from a spatial object


Ideas: Match stems to pixels, etc.

# Is there a relationship between mismatch between LiDAR-measured vegetation height distributions and ground-measured vegetation height distributions as a function of canopy closure?

Compare height distributions 


Load packages and data

```{r load_packages}
# Check working directory
getwd()

# Load our packages
library(raster)
library(rgdal)
library(ggplot2)
library(dplyr)
library(sp)
library(hydroGOF)
```

Load LiDAR CHM raster and remove 0s
```{r load_CHM}
# Load LiDAR CHM raster
site_chm<-raster("../NEONdata/D17-California/SJER/2013/lidar/SJER_lidarCHM.tif")
# look at it
hist(site_chm)

# set 0 values to no data
site_chm[site_chm==0]<-NA
```

Load ground-based measurements
```{r import-field-measurements}
# Import csv field data
site_insitu<-read.csv("../NEONdata/D17-California/SJER/2013/insitu/veg_structure/D17_2013_SJER_vegStr.csv", stringsAsFactors = FALSE) # this is a dataframe
head(site_insitu)
```

Load plot locations
```{r load_plot_locations}

plots<-readOGR("../NEONdata/D17-California/SJER/vector_data/", "SJER_plot_centroids")
plot(plots)
str(plots)
# check to see if there is spatial info
plots
```

Make the plots rectangular ploygons - FAIL!
# maybe add na.rm = TRUE to extract code
```{r make-plots-square-fail, eval=FALSE}
# Define plot boundaries
yPlus <- plot_centroids$northing+20
xPlus <- plot_centroids$easting+20
yMinus <- plot_centroids$northing-20
xMinus <- plot_centroids$easting-20

#Extract the plot ID information
ID=as.character(plot_centroids$Plot_ID) 

#calculate polygon coordinates for each plot centroid. 
square=cbind(xMinus,yPlus, xPlus,yPlus, xPlus,yMinus, xMinus,yMinus,xMinus,yPlus)

#create spatial polygons
site_plots <- SpatialPolygons(
  mapply(
    function(poly, id) {
  xy <- matrix(poly, ncol=2, byrow=TRUE)
  Polygons(list(Polygon(xy)), ID=id)
}, 
split(square, row(square)), ID),
proj4string=CRS(as.character("+proj=utm +zone=11 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))
)

plot(site_plots)
plot(site_chm)
plot(site_plots, add=TRUE)
```

Extract data from LiDAR CHM raster at plots - max and mean heights
```{r extract_CHM_from_plots}
# We're going to extract the mean LiDAR CHM data at the points of the plots
max_height<-extract(site_chm, 
                     plots,
                     buffer=20,
                     fun=max,# we're going to compare max height, so specify fun = max below. You could compare anything or not include a function to get full distribution
                     sp=TRUE, # we want a spatial object
                     stringAsFactors=F # avoid num to characters
                     )
plot(max_height)
max_height@data # this is the dataframe

# We're going to extract the sd LiDAR CHM data at the points of the plots
mean_height<-extract(site_chm, 
                     plots,
                     buffer=20,
                     fun=mean,# we're going to compare max height, so specify fun = max below. You could compare anything or not include a function to get full distribution
                     sp=TRUE, # we want a spatial object
                     stringAsFactors=F # avoid num to characters
                     )
plot(mean_height)
mean_height@data # this is the dataframe
```


Extract mean and tallest tree for each plot 
```{r field-measurements-stats}

# We need to get mean and max tree height for each plot, so extract mean and max values for each plotid
insitu_max_stem_height<-site_insitu %>% 
  group_by(plotid) %>% 
  summarise(max(stemheight))
# take the df and group by plot id, and for each group, give max value
names(insitu_max_stem_height)<-c("plotid", "insitu_maxHt") # rename columns

insitu_mean_stem_height<-site_insitu %>% 
  group_by(plotid) %>% 
  summarise(mean(stemheight))
# take the df and group by plot id, and for each group, give max value
names(insitu_mean_stem_height)<-c("plotid", "insitu_meanHt") # rename columns
```

Merge data in spatial df with mean and max stem height table
```{r merge-data}
# Merge the data together so each point location has LiDAR CHM and insitu measurements
merged.max.height<-merge(max_height@data, insitu_max_stem_height, by.x="Plot_ID", by.y="plotid")

merged.mean.height<-merge(mean_height@data, insitu_mean_stem_height, by.x="Plot_ID", by.y="plotid")


# plot data
ggplot(merged.max.height, aes(x=SJER_lidarCHM, y=insitu_maxHt))+geom_point()
ggplot(merged.mean.height, aes(x=SJER_lidarCHM, y=insitu_meanHt))+geom_point()
```


# quantify error
```{r quantify-errors}
# RMSE for mean and max
rmse(merged.max.height$SJER_lidarCHM, merged.max.height$insitu_maxHt)
# 3.459163
mean(merged.max.height$SJER_lidarCHM) #14.46222
mean(merged.max.height$insitu_maxHt) #13.98333
# mean plot-level mean canopy height is lower in the measured data than LiDAR

rmse(merged.mean.height$SJER_lidarCHM, merged.mean.height$insitu_meanHt) # 3.939722
mean(merged.mean.height$SJER_lidarCHM) #7.559299
mean(merged.mean.height$insitu_meanHt) #5.605737
# mean plot-level mean canopy height is also lower in the measured data than LiDAR. Statistically?


# RMSE is higher with mean canopy height than max canopy height (LiDAR captures max canopy height better than mean canopy heigh), presumably because the data misses understory vegetation
```


# Does RMSE vary ~ plot-level canopy closure?
```{r make_canopy_closure_variable}

# select what 'tall tree' is
# use 3rd quartile of each plot
insitu_stem_height_quartile<-site_insitu %>% 
  group_by(plotid) %>% 
  summarise(tall_tree=quantile(stemheight, probs=0.75))

# merge these
tree_height<-merge(insitu_stem_height_quartile, site_insitu, by="plotid")

#subset just tall trees
# for each tree, if it's as tall as the 3rd quartile for that plot, it's a tall tree. If not, exclude it.
tall_trees<-seq(1, nrow(tree_height), 1)
for(i in 1:nrow(tree_height))
  {
  tall_trees[i]<-ifelse(tree_height$stemheight[i]>=tree_height$tall_tree[i], tree_height$stemheight[i],-2222)
}

# add this to df
tree_height<-cbind(tree_height, tall_trees)
tree_height<-tree_height[tree_height$tall_trees>0,]
```


########BELOW NOT DONE
```{r canopy-closure}

# now we have the tall trees for each plot. We need to sum the canopy diameters for all tall trees for each plot as our measure of canopy closure
tree_height$canopy_area<-pi*(tree_height$maxcanopydiam/2)*(tree_height$canopydiam_90deg/2)

canopy_closure<-aggregate(canopy_area~plotid, data=tree_height, sum)
hist(canopy_closure$canopy_area)

# plot of  ~ canopy closure
# take the df and group by plot id, and for each group, give max value
names(insitu_mean_stem_height)<-c("plotid", "insitu_meanHt") # rename columns
```





# next: explore outliers to see what's going on
# canopy closure is canopy diameter of stems over a certain height



